---
title: "Circa chem"
author: "Stavros Oikonomou"
date: "11/9/2021"
output:
  html_document: default
  word_document: default
---

```{r, warning=FALSE, message=FALSE, include=FALSE}
rm(list = ls(all=TRUE))
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("readxl","dplyr","stringr","NADA","lme4","sjPlot",
              "tableone","knitr","ggplot2","reshape2","tidyr")

ipak(packages)
```


```{r, include=FALSE}
#loading the datasets
pba_cn <- read_excel("./data_files/3PBA-6CN Creatinine CIRCACHEM Final.xlsx", 
                     sheet = "6CN-3PBA", skip=1, 
                     col_names = c("Code", "CN", "PBA","X4","X5","X6","X7"))

df_crea <- read_excel("./data_files/3PBA-6CN Creatinine CIRCACHEM Final.xlsx", 
                      sheet = "Creatinine", col_names = c("Code","Creatinine"),skip=1)

df_hne <- read_excel("./data_files/4HNE CIRCACHEM RESULTS_v2.xlsx", 
                      sheet = "final3", col_names = c("Code","hne"),skip=1)

df_cor <- read_excel("./data_files/Cortisol.xlsx", 
                       col_names = c("cortisol","Code"),skip=1)

df_descr <- read_excel("./data_files/Copy of CIRCA CHEM_Participants Registry_29.12.20_2.xlsx")

```

```{r,include=FALSE}
#setting the cn/pb lod and loq
cn_lod <- 0.075
cn_loq <- 0.226
pb_lod <- 0.049
pb_loq <- 0.146

#setting the cortisol lod and loq
cor_lod <- 0.172
cor_loq <- 0.516
```

```{r, include=FALSE}
#creating the info from Code feature
info_pest <- str_split_fixed(pba_cn$Code, "-",4) %>% data.frame()
info_pest <- info_pest %>% select(-X1) %>%
  rename(record_id=X2, group=X3, sample_number=X4) %>%
  mutate(record_id=as.factor(record_id), 
         group=as.factor(group), 
         sample_number=as.factor(sample_number))

#create record_id, group and sample number from code for cortsol
df_cor <- df_cor %>% separate(
  Code,
  c(NA, "record_id","group","sample_number"))

#bind back to original the splitted information
df_pest <- cbind.data.frame(info_pest,CN=pba_cn$CN,PBA=pba_cn$PBA)

#creating variables for descriptive
df_pest <- df_pest %>% mutate(cnlod=as.factor(ifelse(CN<cn_lod,"Below LOD","Above LOD")),
                              pbalod=as.factor(ifelse(PBA<pb_lod,"Below LOD","Above LOD")),
                              cnlod_TF=ifelse(CN<cn_lod,T,F),
                              cnlodloq=cut(CN, breaks=c(-Inf, cn_lod,cn_loq, Inf),
                                           labels=c("ND","NQ","D&Q")),
                              pblodloq=cut(PBA, breaks=c(-Inf, pb_lod,pb_loq, Inf),
                                           labels=c("ND","NQ","D&Q")))

#creating variables for descriptive
df_cor <- df_cor %>% mutate(corlod=as.factor(ifelse(cortisol<cor_lod,"Below LOD","Above LOD")),
                              corlod_TF=ifelse(cortisol<cor_lod,T,F),
                              corlodloq=cut(cortisol, breaks=c(-Inf, cor_lod,cor_loq, Inf),
                                           labels=c("ND","NQ","D&Q")))

prop.table(table(df_pest$cnlod))
prop.table(table(df_pest$cnlodloq))

prop.table(table(df_pest$pbalod))
prop.table(table(df_pest$pblodloq))

prop.table(table(df_cor$corlod))
prop.table(table(df_cor$corlodloq))

#applying ros function to deal with 6-CN values below LOD. We going to use ros function because with have a big percentage of values below lod.
cnf <- ros(df_pest$CN,df_pest$cnlod_TF, forwardT="log",reverseT="exp", na.action=NULL) %>%
  as.data.frame()

#bind back the new values
df_pest2 <- merge(df_pest,cnf, by.x="CN", by.y="obs") %>%
  select(record_id,group,sample_number,PBA,modeled) %>%
  dplyr::rename(CN=modeled) %>%
  arrange(record_id,group,sample_number)

# For cortisol values < LOD -> LOD/2
df_cor <- df_cor %>% 
  mutate(cor = case_when(cortisol < cor_lod ~ cor_lod/2,
                         TRUE ~ cortisol))

#check PBA and CN for normality
hist(df_pest2$PBA)
hist(df_pest2$CN)

#check cortisol for normality
hist(df_cor$cor)

rm(pba_cn,info_pest, cnf)
```

`r kable(prop.table(table(df_pest$cnlod)),caption="6-CN Above and Below LOD percentage")`

`r kable(prop.table(table(df_pest$cnlodloq)),caption="6-CN Above and Below LOD/LOQ/Above LOQ percentage")`

`r kable(prop.table(table(df_pest$pbalod)),caption="3-PBA Above and Below LOD percentage")`

`r kable(prop.table(table(df_pest$pblodloq)),caption="3-PBA Above and Below LOD/LOQ/Above LOQ percentage")`

`r kable(prop.table(table(df_cor$corlod)),caption="Cortisol Above and Below LOD percentage")`

`r kable(prop.table(table(df_cor$corlodloq)),caption="Cortisol Above and Below LOD/LOQ/Above LOQ percentage")`

```{r, include=FALSE}
#creating the info from Code feature
cr_info <- str_split_fixed(df_crea$Code, "-",4) %>% data.frame()
cr_info <- cr_info %>% select(-X1) %>%
  rename(record_id=X2, group=X3, sample_number=X4) %>%
  mutate(record_id=as.factor(record_id), 
         group=as.factor(group), 
         sample_number=as.factor(sample_number)) %>%
  #we need to change the one digit entries for 1 to 01 in order to match with the PBA/CN dataset
  mutate(record_id=as.factor(ifelse(str_length(record_id)==1,paste0("0",record_id),
                           as.character(record_id))))

#bind back the data
df_cr <- cbind(cr_info,creatinine=df_crea$Creatinine) %>%
  arrange(record_id,group,sample_number)

#checking the normality of creatinine
hist(df_cr$crea)

#counting the number of entries per participant
removing_part <- df_cr %>% group_by(record_id) %>% summarise(n=n())
#participants with not 8 entries need to be excluded
removing_part <- removing_part %>% filter(n<8) %>% pull(record_id)

rm(cr_info, df_crea, df_pest)
```

```{r, include=FALSE, warning=FALSE, message=FALSE}
#creating the info for hne data
hne_info <- str_split_fixed(df_hne$Code, "-",4) %>% data.frame()

hne_info <- hne_info %>% select(-X1) %>%
  rename(record_id=X2, group=X3, sample_number=X4) %>%
  mutate(record_id=as.factor(record_id), 
         group=as.factor(group), 
         sample_number=as.factor(sample_number)) %>%
  #we need to change the one digit entries for 1 to 01 in order to match with the PBA/CN dataset
  mutate(record_id=as.factor(ifelse(str_length(record_id)==1,paste0("0",record_id),
                           as.character(record_id))))



df_hne2 <- cbind(hne_info,hne=df_hne$hne) %>%
  arrange(record_id,group,sample_number) %>%
  mutate(hne=as.numeric(hne))

hist(df_hne2$hne)

rm(hne_info,df_hne, cn_lod, cn_loq, pb_lod, pb_loq)
```


```{r, include=FALSE}
#age and sex data
df_age_sex <- df_descr[1:45,c(1,3,4)] %>% setNames(.,c("record_id","sex","age")) %>%
  mutate(record_id=as.factor(record_id),
         sex=as.factor(ifelse(sex=="F","Female","Male"))) %>%
  mutate(record_id=as.factor(ifelse(str_length(record_id)==1,paste0("0",record_id),
                           as.character(record_id))))

rm(df_descr)
```

```{r, include=FALSE}
#combine PBA/CN/HNE/cortisol and creatinine dataset.
df <- merge(df_pest2,df_cr, all=T, by=c("record_id","group","sample_number")) %>%
  merge(., df_hne2, all=T, by=c("record_id","group","sample_number")) %>%
    merge(., df_cor, all=T, by=c("record_id","group","sample_number")) %>%
  filter(!(record_id %in% removing_part)) %>% #removing non complete data participant
  filter(!(record_id %in% c("43","32"))) %>% #removing participants from compliance filter in previous script
  mutate(record_id=factor(record_id)) %>%
  mutate(cn_adj=CN/creatinine, #adjust measurements with creatinine
         pb_adj=PBA/creatinine,
         hne_adj=hne/creatinine,
         cor_adj=cor/creatinine) %>%
  #creating the phase info
  mutate(phase=case_when(group=="A" & sample_number %in% c("1","2","3","4")~"Morning",
                         group=="A" & sample_number %in% c("5","6","7","8")~"Evening",
                         group=="B" & sample_number %in% c("1","2","3","4")~"Evening",
                         group=="B" & sample_number %in% c("5","6","7","8")~"Morning"))

df_nobl <- df %>% filter(!(sample_number %in% c("1","5"))) # non baseline data

df_bl <- df %>% filter(sample_number %in% c("1","5")) %>% #baseline data
  select(-sample_number)

#combining baseline and treatment data
df_final <- merge(df_nobl,df_bl, by=c("record_id","group","phase"), all.x = T, 
                  suffixes = c("_TR","_BL")) %>%
  mutate(sample_number_order=case_when(sample_number %in% c("2","6")~"1",
                                       sample_number %in% c("3","7")~"2",
                                       sample_number %in% c("4","8")~"3")) %>%
  #log and scale the adjusted cn,pba and hne measurements
  mutate(cn_adj_TR_log_S=scale(log(cn_adj_TR), scale=FALSE),
         cn_adj_BL_log_S=scale(log(cn_adj_BL), scale=FALSE),
         pb_adj_TR_log_S=scale(log(pb_adj_TR), scale=FALSE),
         pb_adj_BL_log_S=scale(log(pb_adj_BL), scale=FALSE),
         hne_adj_TR_log_S=scale(log(hne_adj_TR), scale=FALSE),
         hne_adj_BL_log_S=scale(log(hne_adj_BL), scale=FALSE),
         cor_adj_TR_log_S=scale(log(cor_adj_TR), scale=FALSE),
         cor_adj_BL_log_S=scale(log(cor_adj_BL), scale=FALSE)) %>%
  #join with age and sex data.
  left_join(df_age_sex)

rm(df_bl, df_nobl)

df2 <- na.omit(df)
df2 <- df2[-98,]
```

`r kable(print(CreateTableOne(c("PBA","CN","creatinine","hne","cor"), data=df2,  strata=c("phase"), addOverall=T, testNonNormal = wilcox.test , argsNonNormal = list(paired = TRUE)), nonnormal =c("PBA","CN","creatinine","hne","cor")), "pipe")`

```{r plots, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(df, aes(x=phase,y=CN, fill=phase)) + geom_boxplot() + 
  xlab("Phase") + ylab("CN") + theme_minimal() + ggtitle("6-CN boxplot for morning and evening")

ggplot(df, aes(x=phase,y=PBA, fill=phase)) + geom_boxplot() + 
  xlab("Phase") + ylab("PBA") + theme_minimal() + ggtitle("3-PBA boxplot for morning and evening")

ggplot(df, aes(x=phase,y=creatinine, fill=phase)) + geom_boxplot() + 
  xlab("Phase") + ylab("Creatinine") + theme_minimal() + ggtitle("Creatinine boxplot for morning and evening")

ggplot(df, aes(x=phase,y=hne, fill=phase)) + geom_boxplot() + 
  xlab("Phase") + ylab("4-HNE") + theme_minimal() + ggtitle("4-HNE boxplot for morning and evening")

ggplot(df, aes(x=phase,y=cor, fill=phase)) + geom_boxplot() + 
  xlab("Phase") + ylab("Cortisol") + theme_minimal() + ggtitle("Cortisol boxplot for morning and evening")
```

```{r normality check, echo=FALSE}
#raw data qqplot
qqnorm(df$CN, main="6-CN QQ plot")
qqnorm(df$PBA, main="3-PBA QQ plot")
qqnorm(df$hne, main="4-HNE QQ plot")
qqnorm(df$cor, main="Cortisol QQ plot")


#adj by creatinine qqplot
qqnorm(df$cn_adj, main="6-CN adjusted by creatinine QQ plot")
qqnorm(df$pb_adj, main="3-PBA adjusted by creatinine QQ plot")
qqnorm(df$hne_adj, main="4-HNE adjusted by creatinine QQ plot")
qqnorm(df$cor_adj, main="Cortisol adjusted by creatinine QQ plot")


#log transformed for adjusted data
df <- df %>% mutate(cn_logadj=log(cn_adj),
                    pb_logadj=log(pb_adj),
                    hne_logadj=log(hne_adj),
                    cor_logadj=log(cor_adj))

#qqplot for log transformed adjusted data.
qqnorm(df$cn_logadj, main="6-CN adjusted by creatinine and log transformed QQ plot")
qqnorm(df$pb_logadj, main="3-PBA adjusted by creatinine and log transformed QQ plot")
qqnorm(df$hne_logadj, main="4-HNE adjusted by creatinine and log transformed QQ plot")
qqnorm(df$cor_logadj, main="Cortisol adjusted by creatinine and log transformed QQ plot")

```

```{r plots2, echo=FALSE, message=FALSE, warning=FALSE}
#CN plos
ggplot(df_final, aes(x=record_id,y=cn_adj_TR_log_S,fill=phase)) + geom_boxplot() + 
  xlab("Participants") + ylab("6-CN adjusted,log transformed and centered") + theme_minimal() +
  geom_point(aes(x=record_id, y=cn_adj_BL_log_S, colour=phase), size=2)

ggplot(df_final, aes(x=record_id,y=cn_adj_TR_log_S, fill=phase)) + geom_boxplot() + 
  xlab("Participants") + ylab("6-CN adjusted,log transformed and centered") + theme_minimal() +
  facet_wrap(~phase)

ggplot(df_final, aes(x=phase,y=cn_adj_TR_log_S, fill=phase)) + geom_boxplot() + theme_minimal()
ggplot(df_final, aes(x=phase,y=cn_adj_BL_log_S, fill=phase)) + geom_boxplot() + theme_minimal()


#pb plots
ggplot(df_final, aes(x=record_id,y=pb_adj_TR_log_S,fill=phase)) + geom_boxplot() + 
  xlab("Participants") + ylab("3-PBA adjusted,log transformed and centered") + theme_minimal() +
  geom_point(aes(x=record_id, y=pb_adj_BL_log_S, colour=phase), size=2)

ggplot(df_final, aes(x=record_id,y=pb_adj_TR_log_S,fill=phase)) + geom_boxplot() + 
  xlab("Participants") + ylab("3-PBA adjusted,log transformed and centered") + theme_minimal() +
  facet_wrap(~phase)

ggplot(df_final, aes(x=phase,y=pb_adj_TR_log_S, fill=phase)) + geom_boxplot() + theme_minimal()
ggplot(df_final, aes(x=phase,y=pb_adj_BL_log_S, fill=phase)) + geom_boxplot() + theme_minimal()

#hne
ggplot(df_final, aes(x=record_id,y=hne_adj_TR_log_S,fill=phase)) + geom_boxplot() + 
  xlab("Participants") + ylab("4-HNE adjusted,log transformed and centered") + theme_minimal() +
  geom_point(aes(x=record_id, y=hne_adj_BL_log_S, colour=phase), size=2)

ggplot(df_final, aes(x=record_id,y=hne_adj_TR_log_S,fill=phase)) + geom_boxplot() + 
  xlab("Participants") + ylab("4-HNE adjusted,log transformed and centered") + theme_minimal() +
  facet_wrap(~phase)

ggplot(df_final, aes(x=phase,y=hne_adj_TR_log_S, fill=phase)) + geom_boxplot() + theme_minimal()
ggplot(df_final, aes(x=phase,y=hne_adj_BL_log_S, fill=phase)) + geom_boxplot() + theme_minimal()

```

```{r, include=FALSE}
df_final$phase <- as.factor(df_final$phase)
df_final$pb_adj_BL_log_S <- as.numeric(df_final$pb_adj_BL_log_S)
df_final$pb_adj_TR_log_S <- as.numeric(df_final$pb_adj_TR_log_S)
df_final$cn_adj_BL_log_S <- as.numeric(df_final$cn_adj_BL_log_S)
df_final$cn_adj_TR_log_S <- as.numeric(df_final$cn_adj_TR_log_S)
df_final$hne_adj_BL_log_S <- as.numeric(df_final$hne_adj_BL_log_S)
df_final$hne_adj_TR_log_S <- as.numeric(df_final$hne_adj_TR_log_S)

#creating the day of measurement variable
df_final$days <- ifelse(df_final$sample_number_order=="1",2,
                        ifelse(df_final$sample_number_order=="2",4,7))

#manual correction for some participants who didn't follow the program of urine sample.
df_final[df_final$record_id=="28" & df_final$sample_number=="8","days"] <- 8

df_final[df_final$record_id=="02" & df_final$sample_number=="6","days"] <- 1

df_final[df_final$record_id=="02" & df_final$sample_number=="7","days"] <- 5

df_final2 <- na.omit(df_final)
df_final2 <- df_final2[-73,]
```

`r kable(print(CreateTableOne(c("CN_BL","CN_TR","PBA_BL","PBA_TR","hne_BL","hne_TR","cor_BL","cor_TR","cn_adj_BL","cn_adj_TR","pb_adj_BL", "pb_adj_TR", "hne_adj_BL","hne_adj_TR","cor_adj_BL","cor_adj_TR"), data=df_final2, strata=c("phase"),addOverall=T ,testNonNormal = wilcox.test, argsNonNormal = list(paired = TRUE) ),nonnormal =c("CN_BL","CN_TR","PBA_BL","PBA_TR","hne_BL","hne_TR","cor_BL","cor_TR","cn_adj_BL","cn_adj_TR","pb_adj_BL", "pb_adj_TR", "hne_adj_BL","hne_adj_TR","cor_adj_BL","cor_adj_TR")), "pipe")`

```{r quantiles, include=FALSE}


cn_qo<-round(quantile(df$CN, c(0, .05,.25,.5,.75,.95,1)), digits=2)
pb_qo<-round(quantile(df$PBA, c(0, .05,.25,.5,.75,.95,1)), digits=2)
cr_qo<-round(quantile(df$creatinine, c(0, .05,.25,.5,.75,.95,1)), digits=2)
hne_qo<-round(quantile(na.omit(df$hne), c(0, .05,.25,.5,.75,.95,1)), digits=2)
cor_qo<-round(quantile(na.omit(df$cor), c(0, .05,.25,.5,.75,.95,1)), digits=2)


cnbl_qo<-round(quantile(df_final$CN_BL, c(0, .05,.25,.5,.75,.95,1)), digits=2)
pbbl_qo<-round(quantile(df_final$PBA_BL, c(0, .05,.25,.5,.75,.95,1)), digits=2)
hnebl_qo<-round(quantile(df_final$hne_BL, c(0, .05,.25,.5,.75,.95,1)), digits=2)
corbl_qo<-round(quantile(na.omit(df_final$cor_BL), c(0, .05,.25,.5,.75,.95,1)), digits=2)


cntr_qo<-round(quantile(df_final$CN_TR, c(0, .05,.25,.5,.75,.95,1)), digits=2)
pbtr_qo<-round(quantile(df_final$PBA_TR, c(0, .05,.25,.5,.75,.95,1)), digits=2)
hnetr_qo<-round(quantile(na.omit(df_final$hne_TR), c(0, .05,.25,.5,.75,.95,1)), digits=2)
cortr_qo<-round(quantile(na.omit(df_final$cor_TR), c(0, .05,.25,.5,.75,.95,1)), digits=2)


cnadjbl_qo<-round(quantile(df_final$cn_adj_BL, c(0, .05,.25,.5,.75,.95,1)), digits=2)
pbadjbl_qo<-round(quantile(df_final$pb_adj_BL, c(0, .05,.25,.5,.75,.95,1)), digits=2)
hneadjbl_qo<-round(quantile(df_final$hne_adj_BL, c(0, .05,.25,.5,.75,.95,1)), digits=2)
coradjbl_qo<-round(quantile(na.omit(df_final$cor_adj_BL), c(0, .05,.25,.5,.75,.95,1)), digits=2)


cnadjtr_qo<-round(quantile(df_final$cn_adj_TR, c(0, .05,.25,.5,.75,.95,1)), digits=2)
pbadjtr_qo<-round(quantile(df_final$pb_adj_TR, c(0, .05,.25,.5,.75,.95,1)), digits=2)
hneadjtr_qo <- round(quantile(na.omit(df_final$hne_adj_TR), c(0, .05,.25,.5,.75,.95,1)), digits=2)
coradjtr_qo <- round(quantile(na.omit(df_final$cor_adj_TR), c(0, .05,.25,.5,.75,.95,1)), digits=2)


cnadjbls_qo<-round(quantile(df_final$cn_adj_BL_log_S, c(0, .05,.25,.5,.75,.95,1)), digits=2)
pbadjbls_qo<-round(quantile(df_final$pb_adj_BL_log_S, c(0, .05,.25,.5,.75,.95,1)), digits=2)
hneadjbls_qo <- round(quantile(df_final$hne_adj_BL_log_S, c(0, .05,.25,.5,.75,.95,1)), digits=2)
coradjbls_qo <- round(quantile(na.omit(df_final$cor_adj_BL_log_S), c(0, .05,.25,.5,.75,.95,1)), digits=2)


cnadjtrs_qo<-round(quantile(df_final$cn_adj_TR_log_S, c(0, .05,.25,.5,.75,.95,1)), digits=2)
pbadjtrs_qo<-round(quantile(df_final$pb_adj_TR_log_S, c(0, .05,.25,.5,.75,.95,1)), digits=2)
hneadjtrs_qo <-round(quantile(na.omit(df_final$hne_adj_TR_log_S), c(0, .05,.25,.5,.75,.95,1)), digits=2)
coradjtrs_qo <-round(quantile(na.omit(df_final$cor_adj_TR_log_S), c(0, .05,.25,.5,.75,.95,1)), digits=2)


quantil <- rbind(cn_qo,pb_qo,cr_qo,hne_qo,cor_qo,
                 cnbl_qo,pbbl_qo,hnebl_qo,corbl_qo,
                 cntr_qo,pbtr_qo,hnetr_qo,cortr_qo,
                 cnadjbl_qo,pbadjbl_qo,hneadjbl_qo,coradjbl_qo,
                 cnadjtr_qo,pbadjtr_qo,hneadjtr_qo,coradjtr_qo,
                 cnadjbls_qo,pbadjbls_qo,hneadjbls_qo,coradjbls_qo,
                 cnadjtrs_qo,pbadjtrs_qo,hneadjtrs_qo,coradjtrs_qo)

row.names(quantil) <- c("6-CN","3-PBA","Creatinine","4-HNE","Cortisol",
                        "6-CN Baseline","3-PBA Baseline","4-HNE Baseline","Cortisol Baseline",
                        "6-CN Treatment","3-PBA Treatment","4-HNE Treatment","Cortisol Treatment",
                        "6-CN adjusted Baseline","3-PBA adjusted Baseline","4-HNE adjusted Baseline","Cortisol adjusted Baseline",
                        "6-CN adjusted Treatment","3-PBA adjusted Treatment", "4-HNE adjusted Treatment","Cortisol adjusted Treatment",
                        "6-CN log adjusted Baseline centered","3-PBA log adjusted Baseline centered","4-HNE log adjusted Baseline centered", 
                        "Cortisol log adjusted Baseline centered",
                        "6-CN log adjusted Treatment centered","3-PBA log adjusted Treatment centered","4-HNE log adjusted Treatment centered",
                         "Cortisol log adjusted Treatment centered")
```

### Quantiles

`r kable(quantil)`

```{r, include=FALSE}
#df_final$phase <- relevel(as.factor(df_final$phase), ref="Morning")

cn_model_days <- lmer(cn_adj_TR_log_S ~ phase + cn_adj_BL_log_S + days + sex + age + cor_adj_TR_log_S +
                 + phase*days + (1|record_id),df_final)

tab_model(cn_model_days)

pb_model_days <- lmer(pb_adj_TR_log_S ~ phase + pb_adj_BL_log_S + days + sex + age + cor_adj_TR_log_S +
                 + phase*days + (1|record_id), df_final)

tab_model(pb_model_days)

hne_model_days <- lmer(hne_adj_TR_log_S ~ phase + hne_adj_BL_log_S + days + sex + age
                 + phase*days + (1|record_id), na.omit(df_final))

tab_model(hne_model_days)

hne_pb_model <- lmer(hne_adj_TR_log_S ~ hne_adj_BL_log_S + pb_adj_TR_log_S + days + sex + age
                 + (1|record_id), na.omit(df_final))

tab_model(hne_pb_model)

hne_cn_model <- lmer(hne_adj_TR_log_S ~ hne_adj_BL_log_S + cn_adj_TR_log_S + days + sex + age
                 + (1|record_id), na.omit(df_final))

tab_model(hne_cn_model)
```

### 6-CN model
`r tab_model(cn_model_days)`

### 3-PBA model
`r tab_model(pb_model_days)`

### 4-HNE model
`r tab_model(hne_model_days)`

### 4-HNE ~ 3-PBA model
`r tab_model(hne_pb_model)`

### 4-HNE ~ 6-CN model
`r tab_model(hne_cn_model)`

```{r plots3, echo=FALSE, message=FALSE, warning=FALSE}

df_final_plot <- df_final %>% mutate(days=as.factor(days)) %>% 
  select(days,pb_adj_TR_log_S,pb_adj_BL_log_S, cn_adj_TR_log_S,cn_adj_BL_log_S,
         hne_adj_TR_log_S,hne_adj_BL_log_S,
        cor_adj_TR_log_S,cor_adj_BL_log_S,phase)
levels(df_final_plot$days) <- c(levels(df_final_plot$days),"3","6")

df_final_plot_2 <-data.frame(as.factor(as.character("3")),NA,NA,NA,NA,NA,NA,NA,NA,
                             as.factor(as.character("Evening"))) %>%
                               setNames(names(df_final_plot))

df_final_plot_3 <-data.frame(as.factor(as.character("6")),NA,NA,NA,NA,NA,NA,NA,NA,
                             as.factor(as.character("Evening"))) %>%
                               setNames(names(df_final_plot))

df_final_plot_4 <-data.frame(as.factor(as.character("3")),NA,NA,NA,NA,NA,NA,NA,NA,
                             as.factor(as.character("Morning"))) %>%
                               setNames(names(df_final_plot))

df_final_plot_5 <-data.frame(as.factor(as.character("6")),NA,NA,NA,NA,NA,NA,NA,NA,
                             as.factor(as.character("Morning"))) %>%
                               setNames(names(df_final_plot))

df_final_plot_df <- rbind(df_final_plot, df_final_plot_2,df_final_plot_3,df_final_plot_4,
                          df_final_plot_5)

ggplot(df_final_plot_df, aes(x=as.factor(days),y=pb_adj_TR_log_S, fill=phase)) + geom_boxplot() +
  geom_jitter() + facet_wrap(~phase) + xlab("Days after baseline measurement") +
  ylab("3-PBA log adj cent") + geom_boxplot(aes(x=as.factor("0=BL"),y=pb_adj_BL_log_S),fill=I("Green")) +
  geom_jitter(aes(x=as.factor("0=BL"),y=pb_adj_BL_log_S)) + theme(legend.position = "none")

ggplot(df_final_plot_df, aes(x=as.factor(days),y=cn_adj_TR_log_S, fill=phase)) + geom_boxplot() +
  geom_jitter() + facet_wrap(~phase) + xlab("Days after baseline measurement") +
  ylab("6-CN log adj cent") + geom_boxplot(aes(x=as.factor("0=BL"),y=cn_adj_BL_log_S), fill=I("Green")) +
  geom_jitter(aes(x=as.factor("0=BL"),y=cn_adj_BL_log_S)) + theme(legend.position = "none")

ggplot(df_final_plot_df, aes(x=as.factor(days),y=hne_adj_TR_log_S, fill=phase)) + geom_boxplot() +
  geom_jitter() + facet_wrap(~phase) + xlab("Days after baseline measurement") +
  ylab("4-HNE log adj cent") + 
  geom_boxplot(aes(x=as.factor("0=BL"),y=hne_adj_BL_log_S), fill=I("Green")) +
  geom_jitter(aes(x=as.factor("0=BL"),y=hne_adj_BL_log_S)) + theme(legend.position = "none")

ggplot(df_final_plot_df, aes(x=as.factor(days),y=cor_adj_TR_log_S, fill=phase)) + geom_boxplot() +
  geom_jitter() + facet_wrap(~phase) + xlab("Days after baseline measurement") +
  ylab("Cortisol log adj cent") + 
  geom_boxplot(aes(x=as.factor("0=BL"),y=cor_adj_BL_log_S), fill=I("Green")) +
  geom_jitter(aes(x=as.factor("0=BL"),y=cor_adj_BL_log_S)) + theme(legend.position = "none")
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(df_final, aes(y=hne_adj_TR_log_S,x=cn_adj_TR_log_S, color=sex)) + geom_point( aes(color=sex)) + 
  geom_smooth() + theme_classic() + ylab("4-HNE log centered adjusted") + xlab("6-CN log centered adjusted")

ggplot(df_final, aes(y=hne_adj_TR_log_S,x=pb_adj_TR_log_S,color=sex)) + geom_point(aes(color=sex)) + 
  geom_smooth() + theme_classic() + ylab("4-HNE log centered adjusted") + xlab("3-PBA log centered adjusted")
```

## Sensitivity analysis

exclude participants with compliance <80% in either one of the 2 intervention periods 
```{r, include=FALSE}

# based on previous script 
# firsec = id with consumption below 39 in either one of the 2 weeks
#kable(print(firsec),"pipe")
#[1] "3"  "14" "24" "32" "40" "43" "10" "39"

df_final_sens <- df_final %>% 
  filter(!(record_id %in% c("03","14","24","32","40","43","10","39"))) # exclude 

cn_model_sens <- lmer(cn_adj_TR_log_S ~ phase + cn_adj_BL_log_S + days + sex + age
                 + phase*days + (1|record_id),df_final_sens)

tab_model(cn_model_sens)

pb_model_sens <- lmer(pb_adj_TR_log_S ~ phase + pb_adj_BL_log_S + days + sex + age
                 + phase*days + (1|record_id), df_final_sens)

tab_model(pb_model_sens)

hne_model_sens <- lmer(hne_adj_TR_log_S ~ phase + hne_adj_BL_log_S + days + sex + age
                 + phase*days + (1|record_id), na.omit(df_final_sens))

tab_model(hne_model_sens)

hne_pb_model_sens <- lmer(hne_adj_TR_log_S ~ hne_adj_BL_log_S + pb_adj_TR_log_S + days + sex + age
                 + (1|record_id), na.omit(df_final_sens))

tab_model(hne_pb_model_sens)

hne_cn_model_sens <- lmer(hne_adj_TR_log_S ~ hne_adj_BL_log_S + cn_adj_TR_log_S + days + sex + age
                 + (1|record_id), na.omit(df_final_sens))

tab_model(hne_cn_model_sens)
```

### 6-CN model
`r tab_model(cn_model_sens)`

### 3-PBA model
`r tab_model(pb_model_sens)`

### 4-HNE model
`r tab_model(hne_model_sens)`

### 4-HNE ~ 3-PBA model
`r tab_model(hne_pb_model_sens)`

### 4-HNE ~ 6-CN model
`r tab_model(hne_cn_model_sens)`
